---
title: "Comparisons"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this file, I compare the bechdamic framework with the orignal methods. This
was done to find the parameters that would allow me to replicate the framework
of Calgaro, 2020 in all of the datasets.

```{r, message=FALSE}
library(MicrobiomeBenchmarkData)
library(benchdamic)
library(phyloseq)

library(lefser)
library(DESeq2)
library(ALDEx2)


library(ggplot2)

## Get normalization options for each method in a tibble
source("../R/utils.R")
source("../R/normalization_methods.R")
```

## Data

For this example, I'll use the subgingival vs supragingival plaque datasets.

```{r import data}
grp <- "hmp_body_subsite"
condA <- "subgingival_plaque" # reference
condB <- "supragingival_plaque"

alpha_val <- 0.1

tse = getDataset(x = "HMP_2012_16S_gingival_V35_subset", dryrun = FALSE)[[1]]
colData(tse)[[grp]] <- factor(colData(tse)[[grp]], levels = c(condA, condB))
ps = mia::makePhyloseqFromTreeSummarizedExperiment(tse)
ps
```

```{r extract matrix and conditions}
mat <- otu_table(ps) |> as.data.frame() |> as.matrix()
conditions <- sample_data(ps)[[grp]]
```

## Add normalization factors to sample data

This step is required for benchdamic

```{r normalization factors, message=FALSE}
set_norm <- setNormalizations(
    fun = normalization_methods$fun, method = normalization_methods$method
)

ps <- runNormalizations(
    normalization_list = set_norm, object = ps, verbose = FALSE
)

sample_data(ps)[,startsWith(colnames(sample_data(ps)), "NF.")] |>
    as.data.frame() |>
    head()
```

## DESeq2

```{r deseq2 - original, message=FALSE}
dds <- DESeqDataSet(tse, design = as.formula(paste0("~ ", grp)))
dds <- DESeq2::estimateSizeFactors(object = dds, type = "poscounts")
ddsRes <- DESeq(object = dds, test = "LRT", reduced = ~1, parallel = FALSE)
dds_res <- results(ddsRes)
dds_res |> as.data.frame() |> head()
```


```{r deseq2 - benchdamic, message=FALSE}
deseq2_benchdamic <- set_DESeq2(
    pseudo_count = FALSE, design = as.formula(paste0("~", grp)),
    contrast = c(grp, condB, condA), alpha = alpha_val, norm = "poscounts",
    weights_logical = FALSE, expand = TRUE
)

deseq2_benchdamic_result <- 
    runDA(method_list = deseq2_benchdamic, object = ps, verbose = FALSE)

deseq2_benchdamic_statInfo <- deseq2_benchdamic_result$DESeq2.poscounts$statInfo

head(deseq2_benchdamic_statInfo)
```

The results of both analyses are the same:

```{r}
mapply(dds_res, deseq2_benchdamic_statInfo, FUN = function(x, y) {
    all(x == y)
})
```

## ALDEx2

Directly using the `aldex` function from the ALDEx2 package:

```{r}
aldex_output <- aldex(
    reads = mat, conditions = conditions, mc.samples = 128, test = "t",
    effect = TRUE, include.sample.summary = FALSE, verbose = FALSE,
    denom = "iqlr"
)
```

```{r}
aldex_output |>
    ggplot(aes(effect, -log10(wi.ep))) +
    geom_point()
```


```{r}
aldex_sig <- ifelse(aldex_output$wi.eBH <= 0.1, TRUE, FALSE)

rownames(aldex_output)[aldex_sig]
```

Using benchdamic functions

```{r}
my_aldex2 <- set_ALDEx2(
    pseudo_count = FALSE, conditions = conditions, mc.samples = 128, 
    test = "t", denom = "iqlr", norm = "none", expand = FALSE
)

aldex_benchdamic_output <- runDA(method_list = my_aldex2, object = ps)

```


```{r}
aldex_benchdamic_output$ALDEx2.none.iqlr.t$statInfo |>
    ggplot(aes(effect, -log10(wi.ep))) +
    geom_point()
```

```{r}
aldex_benchdamic_sig <- ifelse(
    aldex_benchdamic_output$ALDEx2.none.iqlr.t$statInfo$wi.eBH <= 0.1, 
    TRUE, FALSE
) 

aldex_benchdamic_output$ALDEx2.none.iqlr.t$statInfo |>
    rownames() |>
    {\(y) y[aldex_benchdamic_sig]}()

```





