---
title: "Comparison of functions to perform CLR"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
if (!'MicrobiomeBenchmarkData' %in% installed.packages()[, 'Package']) {
    BiocManager::install('waldronlab/MicrobiomeBenchmarkData')
}

library(MicrobiomeBenchmarkData)
```

```{r, warning=FALSE}
tse <- getBenchmarkData('Stammler_2016_16S_spikein', dryrun = FALSE)[[1]]
m <- assay(tse, 'counts')
m[1:4, 1:4]
```

# Calculation of geometric mean

There are a few options for calculating the geometric mean in R.

Option 1:

```{r}
gm1 <- apply(m + 1, 2, function(x) exp(mean(log(x))))
gm1
```

Option 2:

```{r}
gm2 <- apply(m + 1, 2, function(x) prod(x)^(1/length(x)))
gm2
```

Note that the second option gives some infinite values. This is because 
the output of `prod` can get really big and cannot be handled by R. The 
`Hotelling:clr` function uses this approach (it can be checked in the function's
definition), so it's better to avoid it. Using the `gmp` package might help (see 
https://www.r-bloggers.com/2019/08/really-large-numbers-in-r/), but the
first approach (`exp(mean(log(x)))`) is better.

# Calculate CLR with other functions

Comparing different functions

```{r}
## Defined function in global environment
my_clr <- function(x) log(x / exp(mean(log(x))))
res <- apply(m + 1, 2, my_clr)

## compositions pkg
res_com <- m |>
    t() |>
    {\(y) compositions::clr(y + 1)}() |>
    t()
class(res_com) <- NULL
attr(res_com, "orig") <- NULL

## mixOmics pkg
res_mix <- m |> 
    t() |>
    {\(y) mixOmics::logratio.transfo(y + 1, logratio = "CLR")}() |>
    t()
class(res_mix) <- NULL # This is needed to remove the "clr" class
```

```{r}
all.equal(res, res_com)
all.equal(res, res_mix)
```

## A simpler test

```{r}
mat <- matrix(0:8, nrow = 3)
mat
```

```{r}
mat_my_clr <- apply(mat + 1, 2, my_clr)
mat_my_clr
```

```{r}
mat_mix <- t(mixOmics::logratio.transfo(t(mat), logratio = "CLR", offset = 1)) |>
    {\(y) `class<-`(y, "matrix")}()
mat_mix
```
