---
title: " HMP_2012_16S_gingival_V35 - subgingival vs supgragingival plaque"
output: 
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=FALSE, warning=FALSE}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(mia)
library(phyloseq)
library(benchdamic)
library(dplyr)
library(purrr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
```

# Data

Import dataset:

```{r import data}
dat_name <- 'HMP_2012_16S_gingival_V35'
conditions_col <- 'body_subsite'
conditions <- c(condB = 'subgingival_plaque', condA = 'supragingival_plaque')

tse <- getBenchmarkData(dat_name, dryrun = FALSE)[[1]]
tse
```

## Exploring some of the data

Let's convert the col_data into a tibble (jsut for ease of handling):

```{r}
col_data <- tse |> 
    colData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column("sample_name") |> 
    as_tibble()
```

Total number of subjects:

```{r}
subjects <- col_data |> 
    pull(subject_id) |> 
    unique()
length(subjects)
```

The number of male and female subjects:

```{r}
col_data |> 
    select(subject_id, gender) |> 
    unique() |> 
    count(gender) |> 
    arrange(-n)
```

Number of subjects per visit number:

```{r}
col_data |> 
    select(subject_id, visit_number) |> 
    unique() |> 
    count(visit_number) |> 
    arrange(-n)
```

Number of subjects per run_center:

```{r}
col_data |> 
    select(subject_id, run_center) |> 
    unique() |> 
    count(run_center) |> 
    arrange(-n)
```

```{r}
sample_names <- vector("list", length(subjects))
names(sample_names) <- subjects
for (i in seq_along(subjects))  {
    current_subject <- subjects[i]
    sub_dat <- col_data |> 
        filter(subject_id == current_subject) |> 
        slice_max(order_by = visit_number, with_ties = TRUE, n = 1)
    if (nrow(sub_dat) < 2) {
        next
    }
    
    lgl_vct <- all(sort(sub_dat[["body_subsite"]]) == conditions)
    if (isFALSE(lgl_vct)) {
        next
    }
    sample_names[[i]] <- sub_dat
}
sample_names <- discard(sample_names, is.null)
col_data_subset <- bind_rows(sample_names)
```

The number of female and male samples is still practically the same

```{r}
col_data_subset |> 
    count(gender)
```

```{r}
selected_samples <- col_data_subset |> 
    pull(sample_name)
tse_subset <- tse[, selected_samples]
tse_subset <- filterTaxa(tse_subset)
tse_subset
```
# Prior information

OTU level:

```{r prior information}
row_data <- as.data.frame(rowData(tse_subset))
prior_info <- row_data[, c('genus', 'taxon_annotation')]
prior_info$taxon_name <- rownames(row_data)
prior_info$new_names <- paste0(prior_info$taxon_name, '|', prior_info$genus)
prior_info <- 
    dplyr::relocate(prior_info, taxon_name, new_names, genus, taxon_annotation)
head(prior_info)
```

Convert to phyloseq

```{r}
ps <- makePhyloseqFromTreeSummarizedExperiment(tse_subset)
sample_data(ps)[[conditions_col]] <- 
    factor(sample_data(ps)[[conditions_col]], levels = conditions)
ps
```

## Differential abundance analysis

Select methods for DA:

```{r}
ps <- runNormalizations(set_norm_list(), ps, verbose = FALSE)
zw <- weights_ZINB(ps, design = conditions_col)
DA_methods <- set_DA_methods_list(conditions_col, conditions)

for (i in seq_along(DA_methods)) {
    if (grepl("Seurat", names(DA_methods)[i])) {
        names(DA_methods[[i]]$contrast) <- NULL
    } else {
        next
    }
}
names(DA_methods)
```

Run all of the differential analysis (DA) methods:

```{r, warning=FALSE}
tim <- system.time({
    DA_output <- vector("list", length(DA_methods))
    for (i in seq_along(DA_output)) {
        message(
            "Running method ", i, ": ", names(DA_methods)[i], " - ", Sys.time()
        )
        DA_output[[i]] <- tryCatch(
            error = function(e) NULL,
            runDA(DA_methods[i], ps, weights = zw, verbose = FALSE) 
        )
    }
    DA_output <- purrr::list_flatten(DA_output, name_spec = "{inner}")
    DA_output <- purrr::discard(DA_output, is.null)
})
tim
```

## Enrichment analysis

Get the column name indicating the direction of the features
(increased or decreased). This is the stats output.

```{r}
direction <- get_direction_cols(DA_output, conditions_col, conditions)
```

## Enrichment (adjP <= 0.1)

Lefse is the strictest method, defining DA with p-values and effect
size thresholds (based on linear discriminant analysis). CLR affects
the magnitud of the effect size, so let's define a new one.

```{r}
DA_output$lefse.TSS$statInfo$abs_score |> hist()
```
```{r}
DA_output$lefse.CLR$statInfo$abs_score |> hist()
```
```{r}
c(
    lefse.TSS = median(DA_output$lefse.TSS$statInfo$abs_score),
    lefse.CLR = median(DA_output$lefse.CLR$statInfo$abs_score)
)
```


Create variables of thresholds:

```{r}
adjThr<- rep(0.1, length(DA_output))
names(adjThr) <- names(DA_output)

esThr <- rep(0, length(DA_output))
names(esThr) <- names(DA_output)
esThr[grep("lefse.TSS", names(esThr))] <- 2
esThr[grep("lefse.CLR", names(esThr))] <- 0.06

slotV <- ifelse(grepl("lefse", names(DA_output)), "statInfo", "pValMat")
colNameV <- ifelse(grepl("lefse", names(DA_output)), "LDA_scores", "adjP")
typeV <- ifelse(grepl("lefse", names(DA_output)), "logfc", "pvalue")
```

Run enrichment:

```{r enrichment}
enrichment <- createEnrichment(
    object = DA_output,
    priorKnowledge = prior_info,
    enrichmentCol = "taxon_annotation",
    namesCol = "new_names",
    slot = slotV, colName = colNameV, type = typeV,
    direction = direction,
    threshold_pvalue = adjThr,
    threshold_logfc = esThr,
    top = NULL, # No top feature selected
    alternative = "greater",
    verbose = FALSE 
)
```

Extract summary of the enrichment analysis:

```{r}
enrichmentSummary <- purrr::map(enrichment,  ~ {
    .x$summaries |> 
        purrr::map(function(x) {
            pos <- which(colnames(x) != "pvalue")
            x |> 
                tibble::rownames_to_column(var = "direction") |> 
                tidyr::pivot_longer(
                    names_to = "annotation", values_to = "n",
                    cols = 2
                )
                
        }) |> 
        dplyr::bind_rows() |> 
        dplyr::relocate(pvalue)
}) |> 
    dplyr::bind_rows(.id = "method") |> 
    dplyr::mutate(
        sig = dplyr::case_when(
            pvalue < 0.05 & pvalue > 0.01 ~ "*",
            pvalue < 0.01 & pvalue > 0.001 ~ "**",
            pvalue < 0.001 ~ "***",
            TRUE ~ ""
        ) 
    ) |> 
    dplyr::mutate(
        direction = dplyr::case_when(
            direction == "DOWN Abundant" ~ "Subgingival",
            direction == "UP Abundant" ~ "Supragingival",
            TRUE ~ direction 
        )
    )


# enrichmentSummary <- enrichmentSummary |> 
#     mutate(
#         method = case_when(
#             grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
#             grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
#             TRUE ~ method
#         )
#     )

```

Create enrichment plot

```{r}
enPlot <- enrichmentSummary |> 
    left_join(get_meth_class(), by = "method") |> 
    mutate(
        direction = factor(
            direction, levels = c("Supragingival", "Subgingival")
        )
    ) |> 
    mutate(
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    ) |> 
    mutate(
        annotation = case_when(
            annotation == "aerobic" ~ "Aerobic",
            annotation == "anaerobic" ~ "Anaerobic",
            annotation == "facultative_anaerobic" ~ "Facultative anaerobic",
            TRUE ~ annotation
        )
    ) |> 
    ggplot(aes(method, n)) +
    geom_col(
        aes(fill = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    geom_text(
        aes(label = sig, color = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    facet_grid(
        direction ~ method_class, scales = "free_x", space = "free"
    ) +
    scale_fill_viridis_d(option = "D", name = "Biological data") +
    scale_color_viridis_d(option = "D", name = "Biological data") +
    labs(
        x = "DA method", y = "Number of DA taxa"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom"
    )
```

# Putative true positives - putative false positives

## Calculate TP - FP ratio (no threshold)

```{r}
positives <- map(1:length(DA_output), .f = function(i) {
    positives <- createPositives(
        object = DA_output[i],
        priorKnowledge = prior_info, 
        enrichmentCol = "taxon_annotation", namesCol = "new_names",
        slot = slotV[i], colName = colNameV[i], type = typeV[i],
        direction = direction[i],
        threshold_pvalue = 1,
        threshold_logfc = 0,
        top = seq.int(from = 0, to = 50, by = 5),
        alternative = "greater",
        verbose = FALSE,
        TP = list(c("DOWN Abundant", "anaerobic"), c("UP Abundant", "aerobic")),
        FP = list(c("DOWN Abundant", "aerobic"), c("UP Abundant", "anaerobic"))
    ) |> 
        left_join(get_meth_class(), by = 'method')
}) |> bind_rows()
# names(positives) <- names(DA_output)


# positives <- createPositives(
#     object = DA_output,
#     priorKnowledge = prior_info, 
#     enrichmentCol = "taxon_annotation", namesCol = "new_names",
#     slot = slotV, colName = colNameV, type = typeV,
#     direction = direction,
#     threshold_pvalue = 1,
#     threshold_logfc = 0,
#     top = seq.int(from = 0, to = 50, by = 5),
#     alternative = "greater",
#     verbose = FALSE,
#     TP = list(c("DOWN Abundant", "anaerobic"), c("UP Abundant", "aerobic")),
#     FP = list(c("DOWN Abundant", "aerobic"), c("UP Abundant", "anaerobic"))
# ) |> 
#     left_join(get_meth_class(), by = 'method')
```

## Plot TP - FP 

```{r, fig.height = 4, fig.width=12}
# names(vec) <- positives$base_method
positives <- positives |> 
    mutate(diff = jitter(TP - FP, amount = 1.5, factor = 2)) |> 
    mutate(
        base_method = case_when(
            grepl("lefse", base_method) ~ sub("lefse", "LEfSe", base_method),
            grepl("wilcox", base_method) ~ sub("wilcox", "Wilcox", base_method),
            TRUE ~ base_method 
        ),
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    )
vec <- positives$color
names(vec) <- positives$base_method
posPlot <- positives |> 
    ggplot(aes(top, diff)) +
    geom_line(
        aes(
            group = method, color = base_method, linetype = norm,
        ),
    ) +
    geom_point(
        aes(
            color = base_method, shape = norm
        ),
    ) +
    facet_wrap(~method_class, nrow = 1) +
    labs(
        x = "Top DA features", y = "TP - FP"
    ) +
    scale_shape(name = "Normalization") +
    scale_linetype(name = "Normalization") +
    scale_color_manual(values = vec, name = "Base method") +
    theme_minimal() +
    theme(legend.position = "bottom")
```

Combine plots:

```{r, fig.height=9, fig.width=10}
pp <- ggarrange(
    plotlist = list(enPlot, posPlot), ncol = 1, heights = c(1.5, 1)
)
pp
```

```{r, eval=TRUE, echo=FALSE}
## Export the figure
ggsave(
    filename = "Figure1.pdf", plot = pp, dpi = 300,
    height = 9, width = 11, 
)
```

# Session Info

```{r}
sessioninfo::session_info()
```
