---
title: "Subgingival vs supgragingival plaque"
subtitle: "HMP_2012_16S_gingival_V35"
author: "Samuel Gamboa"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
    html_document:
        toc: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE
)
```

```{r packages, message=FALSE, warning=FALSE}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(dplyr)
library(purrr)
library(phyloseq)
library(mia)
library(benchdamic)
library(ggplot2)
library(ggpubr)
```

## Data

```{r import data}
dat_name <- 'HMP_2012_16S_gingival_V35'
conditions_col <- 'body_subsite'
conditions <- c(condB = 'subgingival_plaque', condA = 'supragingival_plaque')
tse <- getBenchmarkData(dat_name, dryrun = FALSE)[[1]]
tse
```

```{r filter features}
tse_subset <- filterTaxa(tse, min_ab = 1, min_per = 0.2)
tse_subset
```

```{r samples per body subsite}
col_data <- as_tibble(colData(tse_subset)) 
count(col_data, body_subsite)
```
```{r unique subjects}
length(unique(col_data$subject_id))
```

## Prior information

```{r feature annotations}
row_data <- as.data.frame(rowData(tse_subset))
prior_info <- row_data[, c('genus', 'taxon_annotation')]
prior_info$taxon_name <- rownames(row_data)
prior_info$new_names <- paste0(prior_info$taxon_name, '|', prior_info$genus)
prior_info <- 
    dplyr::relocate(prior_info, taxon_name, new_names, genus, taxon_annotation)
head(prior_info)
```

## Convert to phyloseq

```{r convert to phyloseq}
ps <- convertToPhyloseq(tse_subset)
sample_data(ps)[[conditions_col]] <- 
    factor(sample_data(ps)[[conditions_col]], levels = conditions)
ps
```

## Differential abundance analysis

Set normalization, weight, and DA method options:

```{r norm, weights, methods}
ps <- runNormalizations(set_norm_list(), ps, verbose = FALSE)
zw <- weights_ZINB(ps, design = conditions_col)
DA_methods <- set_DA_methods_list(conditions_col, conditions)
for (i in seq_along(DA_methods)) {
    if (grepl("Seurat", names(DA_methods)[i])) {
        names(DA_methods[[i]]$contrast) <- NULL
    } else {
        next
    }
}
names(DA_methods)
```

Run DA analysis:

```{r run DA, warning=FALSE, message=FALSE}
tim <- system.time({
    DA_output <- imap(DA_methods, ~ {
        message("Running method ", .y, " - ", Sys.time())
        tryCatch(
            error = function(e) NULL,
            runDA(list(.x), ps, weights = zw, verbose = FALSE)
        )
    }) |> 
        list_flatten(name_spec = "{outer}") |> 
        discard(is.null)
    DA_output <- map2(DA_output, names(DA_output), ~ {
        .x$name <- .y
        .x
    })
})
tim
```

## Enrichment analysis

### Define a threshold for LEFSE with CLR

Define threshold variables:

```{r define variables to use in createEnrichment}
direction <- get_direction_cols(DA_output, conditions_col, conditions)

adjThr<- rep(0.1, length(DA_output))
names(adjThr) <- names(DA_output)

esThr <- rep(0, length(DA_output))
names(esThr) <- names(DA_output)
esThr[grep("lefse.TSS", names(esThr))] <- 2
esThr[grep("lefse.CLR", names(esThr))] <-
    median(DA_output$lefse.CLR$statInfo$abs_score)

slotV <- ifelse(grepl("lefse", names(DA_output)), "statInfo", "pValMat")
colNameV <- ifelse(grepl("lefse", names(DA_output)), "LDA_scores", "adjP")
typeV <- ifelse(grepl("lefse", names(DA_output)), "logfc", "pvalue")
```

Run enrichment:

```{r perform enrichment}
enrichment <- createEnrichment(
    object = DA_output,
    priorKnowledge = prior_info,
    enrichmentCol = "taxon_annotation",
    namesCol = "new_names",
    slot = slotV, colName = colNameV, type = typeV,
    direction = direction,
    threshold_pvalue = adjThr,
    threshold_logfc = esThr,
    top = NULL, # No top feature selected
    alternative = "greater",
    verbose = FALSE 
)
```

Extract summary of the enrichment analysis:

```{r enrichment summary}
enrichmentSummary <- purrr::map(enrichment,  ~ {
    .x$summaries |> 
        purrr::map(function(x) {
            x |> 
                tibble::rownames_to_column(var = "direction") |> 
                tidyr::pivot_longer(
                    names_to = "annotation", values_to = "n",
                    cols = 2
                )
                
        }) |> 
        dplyr::bind_rows() |> 
        dplyr::relocate(pvalue)
}) |> 
    dplyr::bind_rows(.id = "method") |> 
    dplyr::mutate(
        sig = dplyr::case_when(
            pvalue < 0.05 & pvalue > 0.01 ~ "*",
            pvalue < 0.01 & pvalue > 0.001 ~ "**",
            pvalue < 0.001 ~ "***",
            TRUE ~ ""
        ) 
    ) |> 
    dplyr::mutate(
        direction = dplyr::case_when(
            direction == "DOWN Abundant" ~ "Subgingival",
            direction == "UP Abundant" ~ "Supragingival",
            TRUE ~ direction 
        )
    )
head(enrichmentSummary)
```

## Plots

### Enrichment plot

```{r enrichment plot}
nn <- unique(enrichmentSummary$annotation)
nn <- nn[!is.na(nn)]
colorPalette <- palette.colors(palette = "Okabe-Ito")[2:(length(nn) + 1)]
enPlot <- enrichmentSummary |> 
    dplyr::left_join(getMethodClass(), by = "method") |> 
    mutate(
        direction = factor(direction, levels = c("Supragingival", "Subgingival")),
        annotation = factor(annotation, levels = nn)
    ) |> 
    mutate(
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    ) |> 
    mutate(
        annotation = case_when(
            annotation == "aerobic" ~ "Aerobic",
            annotation == "anaerobic" ~ "Anaerobic",
            annotation == "facultative_anaerobic" ~ "Facultative anaerobic",
            TRUE ~ annotation
        )
    ) |> 
    ggplot(aes(method, n)) +
    geom_col(
        aes(fill = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    geom_text(
        aes(label = sig, color = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    facet_grid(
        direction ~ method_class, scales = "free_x", space = "free"
    ) +
    scale_fill_manual(values = colorPalette) +
    scale_color_manual(values = colorPalette) +
    labs(
        x = "DA method", y = "Number of DA taxa"
    ) +
    theme_bw() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white")
    )
```

### Putative true positives - putative false positives

Calculate TP - FP ratio (no threshold)

```{r positives object}
positives <- map(1:length(DA_output), .f = function(i) {
    positives <- createPositives(
        object = DA_output[i],
        priorKnowledge = prior_info, 
        enrichmentCol = "taxon_annotation", namesCol = "new_names",
        slot = slotV[i], colName = colNameV[i], type = typeV[i],
        direction = direction[i],
        threshold_pvalue = 1,
        threshold_logfc = 0,
        top = seq.int(from = 0, to = 50, by = 5),
        alternative = "greater",
        verbose = FALSE,
        TP = list(c("DOWN Abundant", "anaerobic"), c("UP Abundant", "aerobic")),
        FP = list(c("DOWN Abundant", "aerobic"), c("UP Abundant", "anaerobic"))
    ) |> 
        dplyr::left_join(getMethodClass(), by = 'method')
}) |> bind_rows()
```

Positives plot:

```{r positives plot, fig.height = 4, fig.width=12}
positives <- positives |> 
    mutate(diff = jitter(TP - FP, amount = 1.5, factor = 2)) |> 
    mutate(
        base_method = case_when(
            grepl("lefse", base_method) ~ sub("lefse", "LEfSe", base_method),
            grepl("wilcox", base_method) ~ sub("wilcox", "Wilcox", base_method),
            TRUE ~ base_method 
        ),
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    )
vec <- positives$color
names(vec) <- positives$base_method
posPlot <- positives |> 
    ggplot(aes(top, diff)) +
    geom_line(
        aes(
            group = method, color = base_method, linetype = norm,
        ),
    ) +
    geom_point(
        aes(
            color = base_method, shape = norm
        ),
    ) +
    facet_wrap(~method_class, nrow = 1) +
    labs(
        x = "Top DA features", y = "TP - FP"
    ) +
    scale_shape(name = "Normalization") +
    scale_linetype(name = "Normalization") +
    scale_color_manual(values = vec, name = "Base method") +
    theme_bw() +
    theme(
        legend.position = "bottom",
        strip.background = element_rect(fill = "white")
    )
```

### Combined plots

```{r combined plots, fig.height=9, fig.width=10}
pp <- ggarrange(
    plotlist = list(enPlot, posPlot), ncol = 1, heights = c(1.5, 1)
)
pp
```

```{r export plots, eval=TRUE, echo=FALSE}
ggsave(
    filename = "Figure1.pdf", plot = pp, dpi = 300,
    height = 9, width = 11, 
)
```

# Session Info

```{r session information}
sessioninfo::session_info()
```
