---
title: " HMP_2012_16S_gingival_V35 - subgingival vs supgragingival plaque"
output: 
    html_document:
        toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup, message=FALSE, warning=FALSE}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(mia)
library(phyloseq)
library(benchdamic)
library(dplyr)
library(purrr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
```

# Data

Import dataset:

```{r import data}
dat_name <- 'HMP_2012_16S_gingival_V35'
conditions_col <- 'body_subsite'
conditions <- c(condB = 'subgingival_plaque', condA = 'supragingival_plaque')

tse <- getBenchmarkData(dat_name, dryrun = FALSE)[[1]]
tse
```

## Exploring some of the data

Let's convert the col_data into a tibble (jsut for ease of handling):

```{r}
col_data <- tse |> 
    colData() |> 
    as.data.frame() |> 
    tibble::rownames_to_column("sample_name") |> 
    as_tibble()
```

Total number of subjects:

```{r}
subjects <- col_data |> 
    pull(subject_id) |> 
    unique()
length(subjects)
```

The number of male and female subjects:

```{r}
col_data |> 
    select(subject_id, gender) |> 
    unique() |> 
    count(gender) |> 
    arrange(-n)
```

Number of subjects per visit number:

```{r}
col_data |> 
    select(subject_id, visit_number) |> 
    unique() |> 
    count(visit_number) |> 
    arrange(-n)
```

Number of subjects per run_center:

```{r}
col_data |> 
    select(subject_id, run_center) |> 
    unique() |> 
    count(run_center) |> 
    arrange(-n)
```

```{r}
sample_names <- vector("list", length(subjects))
names(sample_names) <- subjects
for (i in seq_along(subjects))  {
    current_subject <- subjects[i]
    sub_dat <- col_data |> 
        filter(subject_id == current_subject) |> 
        slice_max(order_by = visit_number, with_ties = TRUE, n = 1)
    if (nrow(sub_dat) < 2) {
        next
    }
    
    lgl_vct <- all(sort(sub_dat[["body_subsite"]]) == conditions)
    if (isFALSE(lgl_vct)) {
        next
    }
    sample_names[[i]] <- sub_dat
}
sample_names <- discard(sample_names, is.null)
col_data_subset <- bind_rows(sample_names)
```

The number of female and male samples is still practically the same

```{r}
col_data_subset |> 
    count(gender)
```

```{r}
selected_samples <- col_data_subset |> 
    pull(sample_name)
tse_subset <- tse[, selected_samples]
tse_subset <- filterTaxa(tse_subset)
tse_subset
```
# Prior information

OTU level:

```{r prior information}
row_data <- as.data.frame(rowData(tse_subset))
prior_info <- row_data[, c('genus', 'taxon_annotation')]
prior_info$taxon_name <- rownames(row_data)
prior_info$new_names <- paste0(prior_info$taxon_name, '|', prior_info$genus)
prior_info <- 
    dplyr::relocate(prior_info, taxon_name, new_names, genus, taxon_annotation)
head(prior_info)
```

Convert to phyloseq

```{r}
ps <- makePhyloseqFromTreeSummarizedExperiment(tse_subset)
sample_data(ps)[[conditions_col]] <- 
    factor(sample_data(ps)[[conditions_col]], levels = conditions)
ps
```

## Differential abundance analysis

Select methods for DA:

```{r}
ps <- runNormalizations(set_norm_list(), ps, verbose = FALSE)
zw <- weights_ZINB(ps, design = conditions_col)
DA_methods <- set_DA_methods_list(conditions_col, conditions)

for (i in seq_along(DA_methods)) {
    if (grepl("Seurat", names(DA_methods)[i])) {
        names(DA_methods[[i]]$contrast) <- NULL
    } else {
        next
    }
}
names(DA_methods)
```

Run all of the differential analysis (DA) methods:

```{r, warning=FALSE}
tim <- system.time({
    DA_output <- vector("list", length(DA_methods))
    for (i in seq_along(DA_output)) {
        message(
            "Running method ", i, ": ", names(DA_methods)[i], " - ", Sys.time()
        )
        DA_output[[i]] <- tryCatch(
            error = function(e) NULL,
            runDA(DA_methods[i], ps, weights = zw, verbose = FALSE) 
        )
    }
    DA_output <- purrr::list_flatten(DA_output, name_spec = "{inner}")
    DA_output <- purrr::discard(DA_output, is.null)
})
tim
```

## Enrichment analysis

Get the column name indicating the direction of the features
(increased or decreased). This is the stats output.

```{r}
direction <- get_direction_cols(DA_output, conditions_col, conditions)
```

## Enrichment (adjP <= 0.1)

```{r enrichment}
enrichment <- createEnrichment(
    object = DA_output,
    priorKnowledge = prior_info,
    enrichmentCol = "taxon_annotation",
    namesCol = "new_names",
    slot = "pValMat", colName = "adjP", type = "pvalue",
    direction = direction,
    threshold_pvalue = 0.1,
    threshold_logfc = 0,
    top = NULL, # No top feature selected
    alternative = "greater",
    verbose = FALSE 
)
```

## Plot enrichment

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
enrich_plot <- plot_enrichment(
    enrichment = enrichment, 
    enrichment_col = "taxon_annotation",
    levels_to_plot = c("aerobic", "anaerobic", "facultative_anaerobic"),
    conditions = conditions
)
p <- plot_enrichment_2(
    enrich_plot, 
    dir = c(up = 'Sup Plq', down = 'Sub Plq')
) +
    theme(
        axis.title = element_text(size = 17),
        axis.text = element_text(size = 15),
        legend.text = element_text(size = 13),
        strip.text = element_text(size = 17)
    )
p
```

# Putative true positives - putative false positives

## Calculate TP - FP ratio (no threshold)

```{r}
positives <- createPositives(
    # object = DA_output, 
    object = DA_output,
    priorKnowledge = prior_info, 
    enrichmentCol = "taxon_annotation", namesCol = "new_names",
    slot = "pValMat", colName = "adjP", type = "pvalue",
    # direction = direction,
    direction = direction,
    threshold_pvalue = 1,
    threshold_logfc = 0,
    top = seq.int(from = 0, to = 50, by = 5),
    alternative = "greater",
    verbose = FALSE,
    TP = list(c("DOWN Abundant", "anaerobic"), c("UP Abundant", "aerobic")),
    FP = list(c("DOWN Abundant", "aerobic"), c("UP Abundant", "anaerobic"))
) |> 
    left_join(get_meth_class(), by = 'method')
```

## Plot TP - FP 

```{r, fig.height = 10, fig.width=15}
positive_plots <- plot_positives(positives) |> 
    map( ~ {
        .x +
            theme(
                axis.title = element_text(size = 17),
                axis.text = element_text(size = 15),
                legend.text = element_text(size = 13),
                strip.text = element_text(size = 17)
            )
    })
k <- grid.arrange(grobs = positive_plots, ncol = 3)
```


```{r, fig.width=15, fig.height=15, warning=FALSE}
figure1 <- ggarrange(p, k, ncol = 1, labels = c("a)", "b)"), heights = c(9, 10))
figure1
```

```{r, eval=TRUE, echo=FALSE}
## Export the figure
ggsave(
    filename = "Figure1.pdf", plot = figure1, dpi = 300,
    width = 15, height = 15
)
```

# Session Info

```{r}
sessioninfo::session_info()
```
