---
title: "Analysis of the HMP_2012_16S_gingival_V35_subset dataset"
output: 
    html_document:
        toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE}
library(MicrobiomeBenchmarkData)
library(MicrobiomeBenchmarkDataAnalyses)
library(phyloseq)
library(mia)
library(benchdamic)
library(purrr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(randomcoloR)
```

# Import data

Dataset

```{r}
tse <- getDataset("HMP_2012_16S_gingival_V35_subset", dryrun = FALSE)[[1]]

grp <- "hmp_body_subsite"

## subgingival will be used as 'control'
conditions <- c(condB = "subgingival_plaque", condA = "supragingival_plaque")
```

# Analysis at the OTU level

Get prior knowledge (biological truth)

```{r}
row_data <- as.data.frame(rowData(tse))
taxaNames <- rownames(row_data)
newNames <- paste0(taxaNames, "|", row_data$GENUS) 
priorInfo <- data.frame(
    taxaNames = taxaNames,
    genus = row_data$GENUS,
    newNames = newNames,
    type = row_data$BIOSIS 
)

rownames(priorInfo) <- taxaNames
head(priorInfo)
```

Convert to phyloseq

```{r}
ps <- makePhyloseqFromTreeSummarizedExperiment(tse)

## This step places the first condition as reference
sample_data(ps)[[grp]] <- factor(
    sample_data(ps)[[grp]], levels = conditions
)
ps
```

## Add normalization factors (benchdamic)

```{r}
norm_pars <- tibble::tribble(
    ~ fun, ~ method,
    "norm_edgeR", "TMM",
    "norm_edgeR", "none",
    "norm_DESeq2", "poscounts",
    "norm_CSS", "median",
    "norm_CSS", "default"
)

norms <- setNormalizations(fun = norm_pars$fun, method = norm_pars$method)

ps <- 
    runNormalizations(normalization_list = norms, object = ps, verbose = FALSE)

colnames(sample_data(ps))
```

## Estimate matrix of weights

These weights are only used by DESeq2, edgeR, and limma-voom to make them
compatible with scRNA-seq analysis.

```{r}
zinbWeights <- weights_ZINB(object = ps, design = grp)
```


## Run DA methods

### DESeq2

```{r}
output_deseq2 <- DA_DESeq2(
    object = ps, pseudo_count = FALSE, 
    design = as.formula(paste0("~", grp)),
    contrast = c(grp, conditions["condA"], conditions["condB"]), 
    norm = "poscounts", 
    verbose = FALSE
)

output_deseq2_zinbweights <- DA_DESeq2(
    object = ps, pseudo_count = FALSE,
    design = as.formula(paste0("~", grp)),
    contrast = c(grp, conditions["condA"], conditions["condB"]), 
    norm = "poscounts", weights = zinbWeights,
    verbose = FALSE
)
```

### edgeR

```{r}
output_edger <- DA_edgeR(
    object = ps, pseudo_count = FALSE, group_name = grp,
    design = as.formula(paste0("~", grp)), norm = "TMM",
    verbose = FALSE
)

output_edger_zinbweihgts <- DA_edgeR(
    object = ps, pseudo_count = FALSE, group_name = grp,
    design = as.formula(paste0("~", grp)), norm = "TMM",
    weights = zinbWeights,
    verbose = FALSE
)
```

### limma

```{r}
output_limma <- DA_limma(
    object = ps, pseudo_count = FALSE, design = grp, norm = "TMM",
    verbose = FALSE,
)

output_limma_zinbweights <- DA_limma(
   object = ps, pseudo_count = FALSE, design = grp, norm = "TMM" ,
   weights = zinbWeights, verbose = FALSE 
)
```

### ALDEx2

```{r}
output_aldex2_t <- DA_ALDEx2(
    object = ps, pseudo_count = FALSE, conditions = grp,
    test = "t", denom = "iqlr", norm = "none",
    verbose = FALSE
)

output_aldex2_wilcox <- DA_ALDEx2(
    object = ps, pseudo_count = FALSE, conditions = grp,
    test = "wilcox", denom = "iqlr", norm = "none",
    verbose = FALSE
)

```

### MetagenomeSeq

```{r}
output_metagenomeseq_default <- DA_metagenomeSeq(
    object = ps, design = grp, coef = 2, norm = "CSSdefault",
    verbose = FALSE
)

output_metagenomeseq_median <- DA_metagenomeSeq(
    object = ps, design = grp, coef = 2, norm = "CSSmedian",
    verbose = FALSE
)

```

### Corncob

```{r}
output_corncob <- DA_corncob(
    object = ps, pseudo_count = FALSE, formula = as.formula(paste0("~", grp)),
    phi.formula = as.formula(paste0("~", grp)),
    formula_null = ~ 1,
    phi.formula_null = as.formula(paste0("~", grp)),
    test = "Wald",
    coefficient = paste0(grp, conditions["condA"]),
    norm = "none",
    verbose = FALSE
)
```


### MAST

```{r}
output_mast <- DA_MAST(
    object = ps, pseudo_count = FALSE, 
    rescale = "median",
    design = as.formula(paste0("~", grp)),
    coefficient = paste0(grp, conditions["condA"]),
    norm = "none",
    verbose = FALSE
)
```

### Seurat

```{r}
output_seurat <- DA_Seurat(
    object = ps, test.use = "wilcox",
    contrast = c(grp, conditions["condA"], conditions["condB"]),
    norm = "none",
    verbose = FALSE
)
```

### Wilcox

```{r, warning=FALSE}
output_wilcox_none <- DA_wilcox(
    object = ps, norm = 'none', design = grp, denom = conditions["condB"]
)

output_wilcox_clr <- DA_wilcox(
    object = ps, norm = 'CLR', design = grp, denom = conditions["condB"]
)

output_wilcox_tss <- DA_wilcox(
    object = ps, norm = 'TSS', design = grp, denom = conditions["condB"]
)
```

### ZINQ

```{r, message=FALSE, warning=FALSE}
output_zinq <- zinq(object = ps, grp = grp, conditions[2])
```


### ANCOMBC

```{r}
output_ancombc <- ancombc(object = ps, formula = grp, group = grp)
```

### LEFSE

```{r}
output_lefse_none <- DA_lefse(
    object = ps, grp = grp, ref = conditions[["condB"]],
    kruskal.threshold = 1,
    wilcox.threshold = 1,
    lda.threshold = 0,
    norm = 'none'
)

output_lefse_clr <- DA_lefse(
    object = ps, grp = grp, ref = conditions[["condB"]],
    kruskal.threshold = 1,
    wilcox.threshold = 1,
    lda.threshold = 0,
    norm = "CLR"
)
```


## Combine DA results

```{r}
DA_results <- list(
    output_deseq2,
    output_deseq2_zinbweights,
    output_edger,
    output_edger_zinbweihgts,
    output_limma,
    output_limma_zinbweights,
    output_aldex2_t,
    output_aldex2_wilcox,
    output_metagenomeseq_median,
    output_metagenomeseq_default,
    output_corncob,
    output_mast,
    output_seurat,
    output_wilcox_none,
    output_wilcox_clr,
    output_wilcox_tss,
    output_zinq,
    output_ancombc,
    output_lefse_none,
    output_lefse_clr
)
names(DA_results) <- map_chr(DA_results, ~ .x$name)
str(DA_results, max.level = 1)

```

## Enrichment

### Direction

```{r}
direction <- c(
    output_deseq2 = "log2FoldChange",
    output_deseq2_zinbweights = "log2FoldChange",
    output_edgeR = "logFC",
    output_edgeR_zinbweihgts = "logFC",
    output_limma = "logFC",
    output_limma_zinbweights = "logFC",
    output_aldex2_t = "effect",
    output_aldex2_wilcox = "effect",
    output_metagenomeseq_default = paste0(grp, conditions["condA"]),
    output_metagenomeseq_median = paste0(grp, conditions["condA"]),
    output_corncob = "Estimate",
    output_mast = "logFC",
    output_seurat = "avg_log2FC",
    output_wilcox_none = "log2FoldChange",
    output_wilcox_clr = "log2FoldChange",
    output_wilcox_tss = "log2FoldChange",
    output_zinq = "log2FoldChange",
    output_ancombc = "logFC",
    output_lefse_none = "LDA_scores",
    output_lefse_clr = "LDA_scores"
)

```

### Calculate enrichment

```{r}
enrichment <- createEnrichment(
    object = DA_results,
    priorKnowledge = priorInfo,
    enrichmentCol = "type",
    namesCol = "newNames",
    slot = "pValMat", colName = "rawP", type = "pvalue",
    direction = direction,
    threshold_pvalue = 0.05,
    threshold_logfc = 0,
    top = NULL,
    alternative = "greater",
    verbose = FALSE 
)

```


### Plot enrichment

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
enrich_plot <- plot_enrichment(
    enrichment = enrichment, 
    enrichment_col = "type",
    levels_to_plot = c("Aerobic", "Anaerobic", "F Anaerobic"),
    conditions = conditions
)
enrich_plot

```

## Putative positives ratio

### Create table of putative positives

```{r}
positives <- createPositives(
    object = DA_results, 
    priorKnowledge = priorInfo, enrichmentCol = "type", namesCol = "newNames",
    slot = "pValMat", colName = "rawP", type = "pvalue",
    direction = direction,
    threshold_pvalue = 0.05,
    threshold_logfc = 0,
    top = seq.int(from = 0, to = 50, by = 5),
    alternative = "greater",
    verbose = FALSE,
    TP = list(c("DOWN Abundant", "Anaerobic"), c("UP Abundant", "Aerobic")),
    FP = list(c("DOWN Abundant", "Aerobic"), c("UP Abundant", "Anaerobic"))
)
```


### Plot table of putative positives

```{r}
# pos_plot <- plotPositives(positives)
# pos_plot
```


```{r}
meth_class <- method_classification()
positives2 <- left_join(meth_class, positives, by = 'method') %>% 
    relocate(method_class)
# positives2 <- positives2 %>% 
#     mutate(
#         base_method = sub("\\..*", "", method)
#     )
```


```{r, fig.height=6, fig.width=10}

# n_meth <- length(unique(positives2$method))
# 
# p1 <- positives2 %>%    
#     mutate(
#         method = paste0(method_class, "--", method)
#     ) %>% 
#     ggplot(
#         aes(x = top, y = TP - FP)
#     ) +
#     geom_path(
#         aes(linetype = method_class, color = method)
#     ) +
#     geom_point(
#         aes(shape = method_class, color = method)
#     ) +
#     labs(x = "Top") +
#     scale_color_manual(values = distinctColorPalette(n_meth)) +
#     scale_linetype_discrete(guide = "none") +
#     scale_shape_discrete(guide = 'none') +
#     facet_wrap(~method_class) +
#     theme_bw() +
#     theme(
#         legend.title = element_blank()
#     )
# 
# p1

```



```{r}

max_top <- max(positives2$TP - positives2$FP)
min_top <- min(positives2$TP - positives2$FP)

x <- split(positives2, positives2$method_class)
df <- x[[2]]
df <- df %>% 
    mutate(
        across(.cols = where(is.character), .fns = ~ forcats::fct_inorder(.x)),
        shape = forcats::fct_inorder(as.character(shape))
        )
        

(
p1 <- df %>%
        ggplot2::ggplot(ggplot2::aes(x = top, y = TP - FP)) +
        ggplot2::geom_path(
            ggplot2::aes(color = color, group = method, linetype = norm)
        ) +
        ggplot2::geom_point(
            ggplot2::aes(color = color, shape = norm),
            size = 3
        ) +
        ggplot2::facet_wrap(.~ method_class) +
        ggplot2::scale_y_continuous(
            limits = c(min_top - 3, max_top + 3)
        ) +
        ggplot2::scale_color_manual(
            values = levels(df$color), labels = levels(df$base_method)
            
        ) +
        ggplot2::scale_shape_manual(
            values = as.integer(levels(df$shape)), labels = levels(df$norm)
        ) +
        ggplot2::scale_linetype_manual(
            values = as.integer(levels(df$shape)), labels = levels(df$norm)
        ) +
        ggplot2::labs(
            x = 'Top', y = 'TP - FP'
        ) +
        ggplot2::theme_bw() +
        ggplot2::theme(
            legend.position = c(0.01, 0.97),
            legend.justification = c("left", "top"),
            legend.title = ggplot2::element_blank(),
            legend.box.background = element_rect(color = "white", fill = 'white'
            )
        ) 
) 

```




```{r, fig.height = 12, fig.width=10}
plots <- plot_positives(positives2)

# first_plot <- plots[1] 
# other_plots <- plots[2:length(plots)]
# other_plots <- lapply(other_plots, function(x) {
#     x <- x +
#         theme(
#             axis.text.y = element_blank(),
#             axis.title.y = element_blank(),
#             axis.ticks.y = element_blank()
#         )
#     x
# })
# 
# plots2 <- c(first_plot, other_plots)

grid.arrange(grobs = plots2, ncol = 2)


```


