---
title: "HMP_2012_16S_gingival_V35"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(mia)
library(phyloseq)
library(magrittr)
library(benchdamic)

library(dplyr)
library(ggplot2)
library(gridExtra)
```

# Data

Import dataset:

```{r import data}
dat_name <- 'HMP_2012_16S_gingival_V35'
x <- getBenchmarkData(dat_name, dryrun = FALSE)[[1]]
x
```
Filter data

+ Subjects from the same run center and the first visit.

```{r}
col_data <- as.data.frame(colData(x))
col_data_sub <-
    dplyr::filter(col_data, run_center == 'WUGC', visit_number == 1)
col_data_split <- split(col_data_sub, factor(col_data_sub$body_subsite))
subjects <- intersect(
    col_data_split$subgingival_plaque$subject_id,
    col_data_split$supragingival_plaque$subject_id
)

male_subjects <- dplyr::filter(
    col_data_sub,
    subject_id %in% subjects, gender == 'male'
) %>% 
    dplyr::pull(subject_id) %>% 
    unique()

female_subjects <- dplyr::filter(
    col_data_sub,
    subject_id %in% subjects, gender == 'female'
) %>% 
    dplyr::pull(subject_id) %>% 
    unique()
n <- min(length(male_subjects), length(female_subjects))
set.seed(1234)
select_subjects <- c(sample(male_subjects, n), sample(female_subjects, n))
select_samples <- rownames(col_data_sub)[
    col_data_sub$subject_id %in% select_subjects
]
x_sub <- x[,select_samples]
x_sub <- x_sub[rowSums(assay(x_sub)) > 0, ]
x_sub
```

Summarize by ranks and retrieve data at the genus level:

```{r}
by_ranks <- splitByRanks(x_sub)
x_genus <- by_ranks$genus
x_genus
```

# Prior information

OTU level:

```{r prior information}
row_data <- as.data.frame(rowData(x_sub))
prior_info <- row_data[, c('genus', 'taxon_annotation')]
prior_info$taxon_name <- rownames(row_data)
prior_info$new_names <- paste0(prior_info$taxon_name, '|', prior_info$genus)
prior_info <- 
    dplyr::relocate(prior_info, taxon_name, new_names, genus, taxon_annotation)
head(prior_info)
```

Genus level

```{r}
row_data_genus <- as.data.frame(rowData(x_genus))
prior_info_genus <- row_data_genus[,c('genus', 'taxon_annotation')]
head(prior_info_genus)
```

# Benchmarking at the OTU level

Define some variables

```{r}
conditions_col <- 'body_subsite'
conditions <- c(condB = 'subgingival_plaque', condA = 'supragingival_plaque')
```

Convert to phyloseq

```{r}
ps <- makePhyloseqFromTreeSummarizedExperiment(x_sub)
ps <- filter_phyloseq(ps)
sample_data(ps)[[conditions_col]] <- 
    factor(sample_data(ps)[[conditions_col]], levels = conditions)
ps
```

## Differential abundance analysis

```{r}
## This can take a while. It's better to save the results and import them back
## later
DA_output_file <-  paste0('./', dat_name, '.rds')
DA_output_flie_2 <- paste0('../../misc/articles_data/', dat_name, '.rds')

if (file.exists(DA_output_file)) {
    message('DA results exist in directory. Importing...')
    DA_output <- readRDS(DA_output_file)
} else if (file.exists(DA_output_flie_2)) {
    message('DA results exist in misc. Importing')
    DA_output <- readRDS(DA_output_flie_2)
} else if (!file.exists(DA_output_file)) {
    message('Performing DA results.')
    DA_output <- run_DA(ps, conditions_col, conditions)
    saveRDS(DA_output, DA_output_file) 
}


# DA_output <- run_DA(ps, conditions_col, conditions)
```


```{r}
remove_output <- c(
    "metagenomeSeq.CSSdefault", "wilcox.none", "ZINQ.none.MinP", 
    "ZINQ.TSS.MinP", "ZINQ.CLR.MinP", "ZINQ.none.Cauchy", "lefse.none", 
    "ALDEx2.none.iqlr.t"
)

DA_output <- DA_output[!names(DA_output) %in% remove_output]
names(DA_output)
```




## Number of DA features

```{r}
lapply(DA_output, function(x) {
    nrow(dplyr::filter(x$pValMat, adjP <= 0.1))
}) %>% 
    unlist() %>% 
    as.data.frame() %>% 
    tibble::rownames_to_column(var = 'method') %>% 
    magrittr::set_colnames(c('method', 'n')) %>% 
    dplyr::arrange(n) %>% 
    dplyr::mutate(method = forcats::fct_inorder(method)) %>% 
    ggplot(aes(method, n)) +
    geom_col() + 
    geom_text(aes(label = n)) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
    
```



## Enrichment analysis


Get direction 

```{r}
direction <- get_direction_cols(DA_output, conditions_col, conditions)
```

## Enrichment (adjP >= 0.1)

```{r}
enrichment <- createEnrichment(
    object = DA_output,
    priorKnowledge = prior_info,
    enrichmentCol = "taxon_annotation",
    namesCol = "new_names",
    slot = "pValMat", colName = "adjP", type = "pvalue",
    direction = direction,
    threshold_pvalue = 0.1,
    threshold_logfc = 0,
    top = NULL, # No top feature selected
    alternative = "greater",
    verbose = FALSE 
)
```

## Plot enrichment

```{r, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
enrich_plot <- plot_enrichment(
    enrichment = enrichment, 
    enrichment_col = "taxon_annotation",
    levels_to_plot = c("aerobic", "anaerobic", "facultative_naerobic"),
    conditions = conditions
)
enrich_plot

```

# Putative true positives - putative false positives

## Calculate TP - FP ratio (adP <= 0.1)

```{r}
positives <- createPositives(
    object = DA_output, 
    priorKnowledge = prior_info, 
    enrichmentCol = "taxon_annotation", namesCol = "new_names",
    slot = "pValMat", colName = "adjP", type = "pvalue",
    direction = direction,
    threshold_pvalue = 1,
    threshold_logfc = 0,
    top = seq.int(from = 0, to = 50, by = 5),
    alternative = "greater",
    verbose = FALSE,
    TP = list(c("DOWN Abundant", "anaerobic"), c("UP Abundant", "aerobic")),
    FP = list(c("DOWN Abundant", "aerobic"), c("UP Abundant", "anaerobic"))
)
```

## Plot TP - FP 

```{r}
meth_class <- get_meth_class() 
positives2 <- left_join(positives, meth_class, by = 'method') %>% 
    relocate(method_class)
```


```{r, fig.height = 12, fig.width=10}
plots <- plot_positives(positives2)
grid.arrange(grobs = plots, ncol = 2)

```


```{r,eval=FALSE, include=FALSE, echo=FALSE}
ggsave(filename = "HMP_2012_a.png", plot = enrich_plot, width = 8, height = 5)

positives_plots <- grid.arrange(grobs = plots, ncol = 3)
ggsave(filename = "HMP_2012_b.png", plot = positives_plots, width = 10, height = 6)
```

