---
title: "Ravel_2011_16S_BV - Bacterial vaginosis - Whole dataset"
output:
    html_document:
        toc: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup, message=FALSE, warning=FALSE}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(mia)
library(phyloseq)
library(dplyr)
library(benchdamic)
library(purrr)
library(ggplot2)
library(gridExtra)
library(ggpubr)
library(tidySummarizedExperiment)
```

## Data

```{r import data, warning=FALSE, message=FALSE}
dat_name <- 'Ravel_2011_16S_BV'
conditions_col <- 'study_condition'
conditions <- c(condB = 'healthy', condA = 'bacterial_vaginosis')

tse <- getBenchmarkData(dat_name, dryrun = FALSE)[[1]]
col_data <- tse |> 
    colData() |> 
    as.data.frame() |> 
    dplyr::filter(study_condition %in% conditions)
col_data |> 
    summarise(
        .by = c(
            "ethnicity", "nugent_score_category", "study_condition"
        ),
        range = paste0(min(nugent_score), "-", max(nugent_score)),
        n = n()
    ) |> 
    arrange(ethnicity, study_condition, n)
```

```{r subset data}
row_names_list <- col_data |>
    {\(y) split(y, factor(y$ethnicity))}() |>
    {\(y) map(y, ~split(.x, .x$study_condition))}() |>
    unlist(recursive = FALSE) |>
    map(rownames)
min_n <- row_names_list |>
    map_int(length) |>
    min()
# set.seed(4567)
# select_samples <- row_names_list |>
#     {\(y) map(y, ~ sample(.x, min_n, replace = FALSE))}() |>
#     unlist(use.names = FALSE)
select_samples <- unique(unlist(row_names_list))
tse_subset <- tse[, select_samples]

tse_genus <- agglomerateByRank(
    tse_subset, rank = 'genus', na.rm = FALSE, onRankOnly = FALSE 
)

tse_genus <- filterTaxa(tse_genus, min_ab = 1, min_per = 0.2)
rownames(tse_genus) <- editMiaTaxaNames(tse_genus)

colData(tse_genus)$study_condition <- 
    factor(colData(tse_genus)$study_condition, levels = conditions)

tse_genus
```

```{r sample counts}
tse_genus |> 
    colData() |> 
    as_tibble() |> 
        summarise(
        .by = c(
            "ethnicity", "nugent_score_category", "study_condition"
        ),
        range = paste0(min(nugent_score), "-", max(nugent_score)),
        n = n()
    ) |> 
    arrange(ethnicity, study_condition, n)
```

## Prior info (biological annotations)

```{r prior info}
prior_info <- tse_genus |> 
    rowData() |> 
    as.data.frame() |> 
    dplyr::select(genus, taxon_annotation) |> 
    rename(taxon_name = genus) |> 
    mutate(
        taxon_annotation = case_when(
            is.na(taxon_annotation) ~ "Unannotated",
            TRUE ~ taxon_annotation
        )
    )
head(prior_info)
```

## DA analysis

Convert to phyloseq:

```{r convert to phyloseq, warning=FALSE}
ps <- convertToPhyloseq(tse_genus)
sample_data(ps)[[conditions_col]] <- factor(
    sample_data(ps)[[conditions_col]], levels = conditions
)
```

Set method parameters:

```{r method parameters, weights, DA methods, warning=FALSE}
norm_methods <- set_norm_list()

ps <- runNormalizations(norm_methods, ps, verbose = FALSE)
zw <- weights_ZINB(ps, design = conditions_col)
DA_methods <- set_DA_methods_list(conditions_col, conditions)

for (i in seq_along(DA_methods)) {
    if (grepl("Seurat", names(DA_methods)[i])) {
        names(DA_methods[[i]]$contrast) <- NULL
    } else {
        next
    }
}
names(DA_methods)
```

Run DA analysis:

```{r run DA, warning=FALSE, message=FALSE}
tim <- system.time({
    DA_output <- vector("list", length(DA_methods))
    for (i in seq_along(DA_output)) {
        message(
            "Running method ", i, ": ", names(DA_methods)[i], " - ", Sys.time()
        )
        DA_output[[i]] <- tryCatch(
            error = function(e) NULL,
            runDA(DA_methods[i], ps, weights = zw, verbose = FALSE) 
        )
    }
    DA_output <- purrr::list_flatten(DA_output, name_spec = "{inner}")
    DA_output <- purrr::discard(DA_output, is.null)
})
tim
```

## Enrichment

We need a threshold for DA for lefse-CLR
(It can't be the same as when using lefse-TSS):

```{r median lefser}
c(
    lefse.TSS = median(DA_output$lefse.TSS$statInfo$abs_score),
    lefse.CLR = median(DA_output$lefse.CLR$statInfo$abs_score)
)
```

Create some variables for selecting and ranking differentially abundant
features:

```{r variables for enrichment funs}
direction <- get_direction_cols(DA_output, conditions_col, conditions)

## The methods based on lefse have artificial p-values because
## the lefser output doesn't provide such information
adjThr<- rep(0.1, length(DA_output))
names(adjThr) <- names(DA_output)

esThr <- case_when(
    grepl("lefse.TSS", names(DA_output)) ~ 2,
    grepl("lefse.CLR", names(DA_output)) ~ 0.15,
    TRUE ~ 0
) |> 
    set_names(names(DA_output))

## Use effect size for lefser and adjusted p-value for all of the other methods
slotV <- ifelse(grepl("lefse", names(DA_output)), "statInfo", "pValMat")
colNameV <- ifelse(grepl("lefse", names(DA_output)), "LDA_scores", "adjP")
typeV <- ifelse(grepl("lefse", names(DA_output)), "logfc", "pvalue")
```

Create enrichment:

```{r enrichment}
enrichment <- createEnrichment(
    object = DA_output,
    priorKnowledge = prior_info,
    enrichmentCol = "taxon_annotation",
    namesCol = "taxon_name",
    slot = slotV, colName = colNameV, type = typeV,
    direction = direction,
    threshold_pvalue = adjThr,
    threshold_logfc = esThr,
    top = NULL, # No top feature selected
    alternative = "greater",
    verbose = FALSE 
)
```

Create enrichment summary:

```{r enrichment summary}
enrichmentSummary <- purrr::map(enrichment,  ~ {
    .x$summaries |> 
        purrr::map(function(x) {
            x |> 
                tibble::rownames_to_column(var = "direction") |> 
                tidyr::pivot_longer(
                    names_to = "annotation", values_to = "n",
                    cols = 2
                )
        }) |> 
        dplyr::bind_rows() |> 
        dplyr::relocate(pvalue)
}) |> 
    dplyr::bind_rows(.id = "method") |> 
    dplyr::mutate(
        sig = dplyr::case_when(
            pvalue < 0.05 & pvalue > 0.01 ~ "*",
            pvalue < 0.01 & pvalue > 0.001 ~ "**",
            pvalue < 0.001 ~ "***",
            TRUE ~ ""
        ) 
    ) |> 
    dplyr::mutate(
        direction = dplyr::case_when(
            direction == "DOWN Abundant" ~ "HV",
            direction == "UP Abundant" ~ "BV",
            TRUE ~ direction 
        )
    )
```

Create enrichment plot:

```{r enrichment plot}
enPlot <- enrichmentSummary |> 
    left_join(get_meth_class(), by = "method") |> 
    mutate(
        direction = factor(
            direction, levels = c("BV", "HV")
        )
    ) |> 
    filter(annotation != "Unannotated") |> 
    mutate(
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    ) |> 
    mutate(
        annotation = case_when(
            annotation == "bv-associated" ~ "BV-associated",
            annotation == "hv-associated" ~ "HV-associated"
        )
    ) |> 
    ggplot(aes(method, n)) +
    geom_col(
        aes(fill = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    geom_text(
        aes(label = sig, color = annotation),
        position = position_dodge2(width = 0.9)
    ) +
    facet_grid(
        direction ~ method_class, scales = "free_x", space = "free"
    ) +
    scale_fill_viridis_d(option = "D", name = "Biological data") +
    scale_color_viridis_d(option = "D", name = "Biological data") +
    labs(
        x = "DA method", y = "Number of DA taxa"
    ) +
    theme_minimal() +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom"
    )
```

## Plot putative true positves and true negatives ratio

Create 'positives' object. No thresholds were added.

```{r putative positives data}
positives <- map(1:length(DA_output), .f = function(i) {
    positives <- createPositives(
        object = DA_output[i],
        priorKnowledge = prior_info, 
        enrichmentCol = "taxon_annotation", namesCol = "taxon_name",
        slot = slotV[i], colName = colNameV[i], type = typeV[i],
        direction = direction[i],
        threshold_pvalue = 1,
        threshold_logfc = 0,
        top = seq.int(from = 2, to = 20, by = 2),
        alternative = "greater",
        verbose = FALSE,
        TP = list(c("DOWN Abundant", "hv-associated"), c("UP Abundant", "bv-associated")),
        FP = list(c("DOWN Abundant", "bv-associated"), c("UP Abundant", "hv-associated"))
    ) |> 
        left_join(get_meth_class(), by = 'method')
}) |>
    bind_rows() |> 
        mutate(
        base_method = case_when(
            grepl("lefse", base_method) ~ sub("lefse", "LEfSe", base_method),
            grepl("wilcox", base_method) ~ sub("wilcox", "Wilcox", base_method),
            TRUE ~ base_method 
        ),
        method = case_when(
            grepl("lefse", method) ~ sub("lefse", "LEfSe", method),
            grepl("wilcox", method) ~ sub("wilcox", "Wilcox", method),
            TRUE ~ method
        )
    )
```

Create "positives" plot:

```{r plot positves, fig.width=15, fig.height=10}
vec <- positives$color
names(vec) <- positives$base_method

posPlot <- positives |> 
    mutate(diff = TP - FP) |>
    ggplot(aes(top, diff)) +
    geom_line(
        aes(
            group = method, color = base_method, linetype = norm,
        ),
    ) +
    geom_point(
        aes(
            color = base_method, shape = norm
        ),
    ) +
    facet_wrap(~method_class, nrow = 1) +
    labs(
        x = "Top DA features", y = "TP - FP"
    ) +
    scale_shape(name = "Normalization") +
    scale_linetype(name = "Normalization") +
    scale_color_manual(values = vec, name = "Base method") +
    theme_minimal() +
    theme(legend.position = "bottom")
```

Combine enrichment and "positives" plot:

```{r combine enrichment and pos plots, fig.height=9, fig.width=10}
pp <- ggarrange(
    plotlist = list(enPlot, posPlot), ncol = 1, heights = c(1.5, 1)
)
pp
```

```{r save plots, eval=FALSE, echo=FALSE}
ggsave(
    filename = "Figure2.pdf", plot = pp,
    dpi = 300, height = 9, width = 11
)
```

# Perform DA with lefse and Wilcox

```{r manual transformations}
tssFun <- function(x) {
    (x) / sum(x) * 1e6
}

clrFun <- function(x) {
    log(x / exp(mean(log(x))))
}

assay(tse_genus, "TSS") <- apply(assay(tse_genus, "counts"), 2, tssFun)
## PseudoCount only needed for CLR
assay(tse_genus, "CLR") <- apply(assay(tse_genus, "counts") + 1, 2, clrFun)

data <- tse_genus |> 
    as_tibble() |> 
    rename(taxon_name = .feature, sample = .sample) |> 
    mutate(
        taxon_annotation = ifelse(
            is.na(taxon_annotation), 'Unannotated', taxon_annotation
        )
    )
head(data)
```

### Wilcox

Define function:

```{r define calcWilcox}
calcWilcox <- function(dat, val_col, log = FALSE) {
    
    ## Separate components
    taxa <- split(dat, factor(dat$taxon_name))
    taxa_names <- names(taxa)
    taxa_annotations <- data |> 
        dplyr::select(tidyselect::starts_with('taxon')) |> 
        dplyr::distinct()
    
    ## Perform Wilcoxon test 
    pvalues <- vector('double', length(taxa))
    names(pvalues) <- taxa_names 
    formula_chr <- paste0(val_col, ' ~ study_condition')
    for (i in seq_along(pvalues)) {
        df <- taxa[[i]]
        res <- stats::wilcox.test(formula = as.formula(formula_chr), data = df)
        pvalues[[i]] <- res$p.value
    }
    
    ## Adjust P-values
    adj_pvalues <- stats::p.adjust(pvalues, method = 'fdr')
    
    ## Calculate fold change
    log_fold_change <- vector('double', length(taxa))
    lll <- vector('double', length(taxa))
    for (i in seq_along(log_fold_change)) {
        df <- taxa[[i]]
        healthy <- df |> 
            dplyr::filter(study_condition == 'healthy') |> 
            {\(y) y[[val_col]]}()
        bv <- df |> 
            dplyr::filter(study_condition == 'bacterial_vaginosis') |> 
            {\(y) y[[val_col]]}()
        
        bv <- mean(bv)
        healthy <- mean(healthy)
        
        if (log) {
            log_fold_change[i] <- bv - healthy
        } else {
            if (bv >= healthy) { # control is healthy, condition of interest is bv
                 log_fold_change[i] <- log2(bv / healthy)
            } else if (bv < healthy) {
                 log_fold_change[i] <- -log2(healthy / bv)
            }
        }
    }
    
    ## Combine results and annotations
    pval_results <- data.frame(
        taxon_name = taxa_names,
        rawP = pvalues,
        adjP = adj_pvalues,
        logFC = log_fold_change
    )
    
    dplyr::left_join(pval_results, taxa_annotations, by = 'taxon_name')
}
```

Perform statistical test:

```{r run calcWilcox, warning=FALSE}
wilcoxRes <- list(
    wilcox_counts = calcWilcox(data, 'counts'),
    wilcox_relab = calcWilcox(data, 'TSS'),
    wilcox_clr = calcWilcox(data, 'CLR', log = TRUE)
) |> 
    bind_rows(.id = 'method')
```

Filter DA taxa

```{r filter taxa}
wilcox_DA <- wilcoxRes |> 
    dplyr::filter(adjP <= 0.1, abs(logFC) > 0) |> 
    mutate(DA = ifelse(logFC > 0, "OA", "UA"))
```

Plot

```{r plot numbers of taxa}
wilcox_DA |> 
    dplyr::filter(taxon_annotation != 'Unannotated') |> 
    count(method, taxon_annotation, DA) |> 
    # mutate(n = ifelse(DA == 'UA', -n, n)) |> 
    mutate(
        DA = case_when(
            DA == "UA"  ~ "HV",
            DA == "OA" ~ "BV"
        )
    ) |> 
    tidyr::complete(DA, method, taxon_annotation, fill = list(n = 0)) |> 
    mutate(method = sub('wilcox_', '', method)) |> 
    ggplot(aes(method, n)) + 
    geom_col(aes(fill = taxon_annotation), position = 'dodge') +
    geom_hline(yintercept = 0) +
    scale_y_continuous(breaks = \(x) pretty(x)) +
    facet_wrap(. ~ DA) + 
    labs(
        title = 'Wilcoxon test',
        y = 'Number of DA taxa', x = 'Transformation method' 
    ) +
    theme_minimal()
```

Plot the abundances of the taxa that were incorrect

```{r incorrect taxa wilcox}
incorrect_taxa_wilcox_clr <- wilcox_DA |> 
    dplyr::filter(
        method == 'wilcox_clr', DA == 'UA', 
        taxon_annotation == 'bv-associated'
    ) |> 
    pull(taxon_name)
incorrect_taxa_wilcox_clr
```

Let's plot their values for each matrix

```{r get abundance values}
transformations <- c('counts', 'TSS', 'CLR')
l1 <- vector('list', length(transformations))
names(l1) <- transformations
for (i in seq_along(transformations)) {
    mat <- assay(tse_genus, transformations[i])
    l1[[i]] <- mat[incorrect_taxa_wilcox_clr,] |> 
        as.data.frame() |> 
        tibble::rownames_to_column(var = 'taxon_name') |> 
        as_tibble()
    
}

wilcox_raw <- bind_rows(l1, .id = 'transformation') |> 
    {\(y) pivot_longer(
        y, cols = 3:ncol(y), values_to = 'value', names_to = 'sample'
    )}() |> 
    left_join(
        distinct(data[,c('sample', 'study_condition')]), by = 'sample'
    )

head(wilcox_raw)
```

Box plot of incorrect values:

```{r plot abundance values, fig.height=7}
l <- wilcox_raw |> 
    mutate(taxon_name = sub('genus:', '', taxon_name)) |> 
    {\(y) split(y, y$transformation)}()
l$counts$value <- log(l$counts$value + 1)
l$TSS$value <- log(l$TSS$value + 1)
## CLR is already in log scale
wilcox_raw <- reduce(l, bind_rows)

wilcox_genus_plot <- wilcox_raw |> 
    mutate(transformation = factor(
        transformation, levels = c('counts', 'TSS', 'CLR'),
        labels = c('log(counts + 1)', 'log(TSS + 1)', 'CLR')
    )) |> 
    mutate(study_condition = factor(
        study_condition, levels = c('healthy', 'bacterial_vaginosis'),
        labels = c('HV', 'BV')
    )) |> 
    ggplot(aes(taxon_name, value)) + 
    geom_boxplot(aes(color = study_condition), position = position_dodge2(0.9)) +
    # geom_point(
    #     aes(color = study_condition), position = position_dodge2(0.9),
    #     size = 0.1
    # ) +
    facet_wrap(~ transformation, scales = 'free') +
    labs(
        y = 'Abundance values', x = 'Genus'
    ) +
    # scale_color_manual(
    #     values = c('dodgerblue1', 'firebrick1')
    # ) +
    theme_bw() +
    theme(
        panel.grid.major.x = element_blank(),
        legend.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, face = "italic")
        
    )
wilcox_genus_plot
```

```{r save plot, eval=FALSE, echo=FALSE}
ggsave(
    file = "Figure3.pdf", plot = wilcox_genus_plot,
    dpi = 300, width = 9, height = 4
)
```

### Lefse

Define a function for running Lefse:

```{r lefse function}
calcLefse <- function(se, assay) {
    res <- lefser::lefser(
        se, kruskal.threshold = 0.05, wilcox.threshold = 0.05, 
        lda.threshold = 0.15, classCol = 'study_condition', assay = assay
    )
    
    return(res)
}

```

Run lefse

```{r lefse run, warning=FALSE}
taxa_annotations <-
        dplyr::distinct(dplyr::select(data, dplyr::starts_with('taxon')))
lefse_DA <- list(
    lefse_counts = calcLefse(tse_genus, 'counts'),
    lefse_relab = calcLefse(tse_genus, 'TSS'),
    lefse_clr = calcLefse(tse_genus, 'CLR')
)  |> 
    bind_rows(.id = 'method') |> 
    mutate(
        DA = ifelse(scores > 0.15, 'OA', 'UA')
    ) |> 
    rename(taxon_name = 'features') |> 
    left_join(taxa_annotations, by = 'taxon_name')
head(lefse_DA)
```
Plot lefse results:

```{r plot DA}
lefse_DA |> 
    dplyr::filter(taxon_annotation != 'Unannotated') |> 
    count(method, taxon_annotation, DA) |> 
    # mutate(n = ifelse(DA == 'UA', -n, n)) |> 
    mutate(
        DA = case_when(
            DA == "UA" ~ "HV",
            DA == "OA" ~ "BV"
        )
    ) |> 
    mutate(method = sub('lefse_', '', method)) |> 
    tidyr::complete(method, taxon_annotation, DA, fill = list(n = 0)) |> 
    ggplot(aes(method, n)) + 
    geom_col(aes(fill = taxon_annotation), position = 'dodge') +
    geom_hline(yintercept = 0) +
    facet_wrap(. ~ DA) +
    scale_y_continuous(breaks = pretty) +
    labs(
        title = 'LEfSe test',
        y = 'Number of DA taxa', x = 'Transformation/Normalization method' 
    ) +
    theme_minimal()
```

```{r get incorrect taxa lefse}
incorrect_taxa_lefse_clr <- lefse_DA |> 
    dplyr::filter(
        method %in% c('lefse_clr', 'lefse_relab_clr'), DA == 'UA', 
        taxon_annotation == 'bv-associated'
    ) |> 
    pull(taxon_name) |> 
    unique()
incorrect_taxa_lefse_clr ## the same as in wilcox.
```

## Compositions with TSS and CLR

Plot TSS:

```{r}
first_set <- data |> 
    filter(
        nugent_score_category == 'low',
        taxon_annotation == 'hv-associated'
    ) |> 
    arrange(desc(TSS)) |> 
    pull(sample)
second_set <- data |> 
    filter(
        nugent_score_category == 'high',
        taxon_annotation == 'hv-associated'
    ) |> 
    arrange(desc(TSS)) |> 
    pull(sample)
samples_order <- c(first_set, second_set)

p1 <- data |> 
    mutate(
        sample = factor(sample, levels = samples_order),
        nugent_score_category = factor(
            nugent_score_category, levels = c('low', 'high'),
            labels = c('Low Nugent score', 'High Nugent score')
        ),
        taxon_annotation = case_when(
            taxon_annotation == "hv-associated" ~ "Health-associated",
            taxon_annotation == "bv-associated" ~ "BV-associated",
            TRUE ~ taxon_annotation
        ),
        taxon_annotation = factor(
            taxon_annotation, levels = c('Health-associated', 'BV-associated', 'Unannotated')[3:1]
        )
    ) |>
    ggplot(aes(sample, TSS )) +
    geom_col(aes(fill = taxon_annotation)) +
    scale_fill_manual(values = c('gray60', 'firebrick2', 'dodgerblue2')) +
    labs(
        x = "Samples",
        y = "Relative abundance values (TSS)"
    ) +
    facet_wrap(~nugent_score_category, ncol = 2, scales = "free_x") +
    theme_bw() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank()
    )
p1
```

Plot CLR:

```{r}
p2 <- data |> 
    mutate(
        sample = factor(sample, levels = samples_order),
        nugent_score_category = factor(
            nugent_score_category, levels = c('low', 'high'),
            labels = c('Low Nugent score', 'High Nugent score')
        ),
        taxon_annotation = case_when(
            taxon_annotation == "hv-associated" ~ "Health-associated",
            taxon_annotation == "bv-associated" ~ "BV-associated",
            TRUE ~ taxon_annotation
        ),
        taxon_annotation = factor(
            taxon_annotation, levels = c('Health-associated', 'BV-associated', 'Unannotated')[3:1]
        )
    ) |>
    ggplot(aes(sample, exp(CLR))) +
    geom_col(aes(fill = taxon_annotation)) +
    scale_fill_manual(values = c('gray60', 'firebrick2', 'dodgerblue2')) +
    labs(
        x = "Samples",
        y = "Geometric mean normalization (exp(CLR))"
    ) +
    facet_wrap(~nugent_score_category, ncol = 2, scales = "free_x") +
    theme_bw() +
    theme(
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        panel.grid = element_blank()
    )
p2
```

## Distributions

Get Latobacillus relative abundance per sample

```{r}
sample_sizes <- filter(data, taxon_name == 'genus:Lactobacillus') |> 
    select(sample, lact_tss = TSS, lact_clr = CLR)
data_with_lact <- left_join(data, sample_sizes, by = 'sample')
```

Plotting log(CLR) vs log(Relab) of Lactobacillus, Prevotella, Actinomyces, and
Corynebacterium.

```{r, fig.height=5, fig.width=7}
plot_1b <- data_with_lact |> 
    filter(taxon_name == 'genus:Actinomyces') |> 
    mutate(
        study_condition = factor(
            study_condition, levels = c('bacterial_vaginosis', 'healthy'),
            labels = c('BV', 'HV')
        )
    ) |> 
    ggplot(aes(log(TSS + 1), CLR)) +
    geom_point(
        aes(color = study_condition, size = lact_tss), 
        alpha = 0.3
    ) + 
    labs(
        # title = 'Relative abundace vs CLR',
        title = expression(italic('Actinomyces') ~ '(BV-associated)'),
        x = 'log(TSS + 1)'
    ) +
    scale_color_discrete(name = 'Condition') +
    # scale_size(name = 'Lactobacillus Rel. Ab.') +
    scale_size(name = expression(italic('Lactobacillus') ~ 'Rel. Ab.')) +
    theme_bw()

plot_2b <- data_with_lact |> 
    filter(taxon_name == 'genus:Corynebacterium') |> 
    mutate(
        study_condition = factor(
            study_condition, levels = c('bacterial_vaginosis', 'healthy'),
            labels = c('BV', 'HV')
        )
    ) |> 
    ggplot(aes(log(TSS + 1), CLR)) +
    geom_point(
        aes(color = study_condition, size = lact_tss), 
        alpha = 0.3
    ) + 
    labs(
        # title = 'Relative abundace vs CLR',
        title = expression(italic('Corynebacterium') ~ '(BV-associated)'),
        x = 'log(TSS + 1)'
    ) +
    scale_color_discrete(name = 'Condition') +
    # scale_size(name = 'Lactobacillus Rel. Ab.') +
    scale_size(name = expression(italic('Lactobacillus') ~ 'Rel. Ab.')) +
    theme_bw()

plot_3b <- data_with_lact |> 
    filter(taxon_name == 'genus:Prevotella') |> 
    mutate(
        study_condition = factor(
            study_condition, levels = c('bacterial_vaginosis', 'healthy'),
            labels = c('BV', 'HV')
        )
    ) |> 
    ggplot(aes(log(TSS + 1), CLR)) +
    geom_point(
        aes(color = study_condition, size = lact_tss), 
        alpha = 0.3
    ) + 
    labs(
        # title = 'Relative abundace vs CLR',
        title = expression(italic("Prevotella") ~ "(BV-associated)"),
        x = 'log(TSS + 1)'
    ) +
    scale_color_discrete(name = 'Condition') +
    # scale_size(name = 'Lactobacillus Rel. Ab.') +
    scale_size(name = expression(italic('Lactobacillus') ~ 'Rel. Ab.')) +
    theme_bw()

plot_4b <- data_with_lact |> 
    filter(taxon_name == 'genus:Lactobacillus') |> 
    mutate(
        study_condition = factor(
            study_condition, levels = c('bacterial_vaginosis', 'healthy'),
            labels = c('BV', 'HV')
        )
    ) |> 
    ggplot(aes(log(TSS + 1), CLR)) +
    geom_point(
        aes(color = study_condition, size = lact_tss), 
        alpha = 0.3
    ) + 
    labs(
        # title = 'Relative abundace vs CLR',
        # title = 'Lactobacillus (HV-associated)',
        title = expression(italic("Lactobacillus") ~ "(HV-associated)"),
        x = 'log(TSS + 1)'
    ) +
    scale_color_discrete(name = 'Condition') +
    # scale_size(name = 'Lactobacillus Rel. Ab.') +
    scale_size(name = expression(italic('Lactobacillus') ~ 'Rel. Ab.')) +
    theme_bw()

plotsb <- ggpubr::ggarrange(
    plot_4b, plot_3b, plot_1b, plot_2b, align = 'hv',
    common.legend = TRUE, legend = 'bottom', 
    labels = c('a)', 'b)', 'c)', 'd)')
) 
```


```{r, fig.height=6, fig.width=8}
plotsb
```

```{r, echo=FALSE}
ggsave(
    plotsb, filename = 'Figure4.pdf',
    width = 8, height = 6, dpi = 300
)
```

# Session info

```{r}
sessioninfo::session_info()
```
