---
title: "Sammler_2016_16S_spikein - Spike-in bacteria"
subtitle:
    "Comparison of coefficient of variation of spike-in bacteria across samples"
output:
    html_document:
        toc: yes
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(MicrobiomeBenchmarkDataAnalyses)
library(MicrobiomeBenchmarkData)
library(dplyr)
library(tibble)
library(tidyr)
library(biobroom)
library(ggplot2)
library(purrr)
```

# Introduction

The objective of this vignette is to compare the coefficient of variation
(CV) of the three spike-in bacteria in the *Stammler\_2016\_16S\_spikein*
dataset using relative abundance (TSS) and centered-log-ratio transformation
(CLR). However, since CLR is log transformed and TSS is not, a
geometric mean normalization (GMN) method will be used instead of CLR.

*S. ruber* will be used for re-calibrating the abundance data (counts).
This is referred to as SCML. The data will also be normalized with
TSS (TSS) or GMN (GMN). The CV of the abundance data
of the three spike-in bacteria across samples using these two normalization
methods will be calculated and compared.

# Data

```{r, message=FALSE, warning=FALSE}
tse <- getBenchmarkData('Stammler_2016_16S_spikein', dryrun = FALSE)[[1]]
```

## Recalibrate with spike-in Salinibacter ruber

```{r}
spk_bac <- c(
    `S. ruber` = 'AF323500XXXX', 
    `R. radiobacter` = 'AB247615XXXX',
    `A. acidiphilus` = 'AB076660XXXX'
)
counts <- assay(tse, 'counts')
s_ruber <- counts[spk_bac['S. ruber'], ]
size_factor <- s_ruber/mean(s_ruber)
SCML_data <- counts 
for(i in seq(ncol(SCML_data))){
    SCML_data[,i] <- round(SCML_data[,i] / size_factor[i])
}
assay(tse, 'SCML') <- SCML_data
```

## Tranform with TSS (relative abundance) and CLR

```{r}
tss_fun <- function(x) (x + 1) / sum((x + 1))
# tss_fun <- function(x) log((x + 1) / sum((x + 1)))
# tss_fun <- function(x) (x) / sum((x))
gmn_fun <- function(x) (x + 1) / exp(mean(log((x + 1))))
# gnm_fun <- function(x) (x + 1) / prod((x + 1)^(1 / length(x)))
# gmn_fun <- function(x) log((x + 1) / exp(mean(log((x + 1)))))
assay(tse, "TSS") <- apply(assay(tse, 'counts'), 2,  tss_fun)
assay(tse, "GMN") <- apply(assay(tse, 'counts'), 2, gmn_fun) 
```

# Extract data of sipike-in bacteria

```{r}
spk_bac_tse <- tse[spk_bac,]
rownames(spk_bac_tse) <- names(spk_bac)
spk_bac_tse
```

# Get tidy data

```{r, message=FALSE, warning=FALSE}
data <- spk_bac_tse |> 
    assays() |> 
    names() |> 
    map({
        ~ tidy.RangedSummarizedExperiment(spk_bac_tse, assay = .x) |> 
            magrittr::set_colnames(c("taxon", "sample", .x))
    }) |> 
    reduce(.f = \(.x, .y) left_join(.x, .y, by = c("taxon", "sample")))
DT::datatable(data, filter = 'top')
```

# Calculate coefficient of variation

## Define a formula for calculating coefficient of variation

```{r}
get_cv <- function(x) {
    cv <- function(x, n) { sd(x[n]) / abs(mean(x[n])) } 
    boot::boot(x, cv, R = 1000) |>
        broom::tidy() |>
        dplyr::rename(cv = statistic)
} 
```

```{r}
cv_res <- data %>% 
    group_by(taxon) %>% 
    summarize(across(.cols = counts:last_col(), .fns = get_cv)) %>% 
    pivot_longer(
        cols = 2:last_col(), names_to = 'norm', values_to = 'cv_res' 
    ) %>% 
    unnest(cols = 'cv_res')
   
DT::datatable(cv_res, filter = 'top')     
```

Table in wider format:

```{r}
cv_res |> 
    rename(
        Species = taxon, `Normalization method` = norm,
        CV = cv, SE = std.error
    ) |> 
    filter(`Normalization method` %in% c("GMN", "TSS")) |> 
    select(-bias) |> 
    mutate(
        CV = round(CV, 2), SE = round(SE, 2)
    ) |> 
    mutate(
        `Normalization method` = ifelse(
            test = `Normalization method` == "TSS", 
            yes = "Relative abundance",
            no = `Normalization method`
        )
    ) |> 
    DT::datatable(
        extensions = 'Buttons',,
        filter = "top",
        options = list(
        dom = 'Bfrtip',
        buttons = list(
            list(
                extend = 'copy',
                text = 'Copy '
                )
            )
        )
    )
```

# Compare coefficient of variation

```{r, fig.width=8, fig.height=4}
cv_res |>  
    filter(norm != 'counts') |> 
    mutate(
        norm = factor(norm, levels = c(
            'counts', 'SCML', 'TSS',  'GMN'
            )
        )
    ) %>%
    mutate(taxon = forcats::fct_relevel(taxon, 'S. ruber')) |> 
    ggplot(aes(reorder(norm, cv), cv)) +
    geom_point(aes(color = norm), size = 2) + 
    geom_errorbar(
        aes(ymin = cv - std.error, ymax = cv + std.error, color = norm),
        width = 0.4, size = 0.5
    ) +
    scale_color_brewer(type = 'qual', palette = 'Set2') +
    facet_wrap(~taxon) + 
    labs(
        y = 'Coefficient of variation across all samples',
        x = 'Data transformation'
    ) + 
    theme_bw() + 
    theme(
        # axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_blank(),
        strip.text = element_text(face = 'italic'),
        legend.position = 'none'
    )
```

# Conclusion

TSS normalizaion has lower CV values than GMN.
This could indicate that it introduces less bias than GMN,
which is related to CLR.

# Session info

```{r}
sessioninfo::session_info()
```
